---
title: "HW4_IE6600_Sec04_Group04"
author: "Parth Urvishkumar Khamar"
date: "2022-10-11"
output:
  word_document: default
  html_document: default
---
```{r Loading Libraries}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(reshape2)
library(extrafont)
#loadfonts(device = "win")
```

Problem Statement: 
Purpose of this assignment to teach, how to use different data and perform analysis using plots via different libraries and functions.


```{r Loading Dataset}
air_df<-read.csv('airlines_delay.csv', na.strings = "")
head(air_df)
```
# Task 1 
Problem Statement:
Generate a density plot by delay_type

```{r }

delays <- air_df %>%
  summarise(carrier_delay, nas_delay, security_delay,
            weather_delay, late_aircraft_delay) %>%
  drop_na()
delays$ID <- seq.int(nrow(delays)) 
long= melt(delays, id.vars= 'ID')
ggplot(long,aes(x=log10(value),fill=variable))+geom_density(alpha=0.5)+labs(title="Density plot of carrier delays in mins",x="log delay in mins", y = "Density") + labs(fill = "Delay type")
```

Conclusion: After monitering this plot, we can see that major delays are happened because of weather

## Task 2 
# Problem Statement:  
Create correlation plots for arr_flights, arr_del15, arr_cancelled, arr_diverted, arr_delay, carrier_delay,
weather_delay, nas_delay, security_delay and late_aircraft_delay with upper half containing correlation plot and lower triangle with magnitude of correlation with the labels not overlapping


```{r }
fig.width=8
library(corrplot)
library(tidyr)
library(data.table)
air <- setnames(air_df, old = c("arr_flights","arr_del15","arr_cancelled","arr_diverted","arr_delay","carrier_delay","weather_delay","nas_delay","security_delay","late_aircraft_delay"), 
new = c('arr \n flights','arr \n del15','arr \n cancelled','arr \n diverted','arr \n delay','carrier \n delay','weather \n delay','nas \n delay','security \n delay','late \n aircraft \n delay'))

df<- air %>%
  select('arr \n flights', 'arr \n del15','arr \n cancelled',
         'arr \n diverted', 'arr \n delay',
         'carrier \n delay','weather \n delay',
         'nas \n delay', 'security \n delay', 'late \n aircraft \n delay') %>%
  drop_na()
M<- cor(df)
corrplot.mixed(M, lower.col="black", number.cex=1, tl.cex=1, cl.cex=1) 
```

Conclusion: From the correlation matrix we see that all the delays are in positive value that everyone 


```{r}
wages<- read.csv('wages_jobs.csv', na.strings = "")
head(wages)
colnames(wages)
```

#Task 3
Problem Statement:
generate a heat map to plot the gender difference with occupation over the year

```{r fig, fig.width=10}
library(scales)
d<- wages %>%
  select(Gender, Year, PUMS.Occupation, Total.Population )%>%
  group_by(PUMS.Occupation, Gender) %>%
  drop_na()
f<- filter(d,Gender=='Female')
m<-filter(d, Gender=='Male')
df2<- merge(f, m, by=c("Year","PUMS.Occupation"))
df3<-mutate(df2, avg_num_diff= Total.Population.y-Total.Population.x)
plt1 <- df3 %>%
  select(Year, PUMS.Occupation, avg_num_diff) %>%
  ggplot( aes(x=Year,y= PUMS.Occupation)) + geom_tile(aes(fill= avg_num_diff), alpha=0.9) + coord_fixed() +
  labs(title="Difference between number of Male and Female Employees",x="Year", y = "Occupation") +
  theme(plot.title = element_text(size=9)) + labs(fill = "Difference") 
plt1
```

#conclusion:
The above heat map shows that more females are working as a flight attendents whereas more males are working as a aircraft pilots & flight engineers.


## Task 4
# Problem Statement:
Create an Alluvial chart with number of male and female workers from each  different PUMS. Occupations in each year.

```{r fig, fig.width=10}
library(tidyverse)
library(ggplot2)
library(ggalluvial)
plt2 <- ggplot(data = d,
               aes(axis1 = PUMS.Occupation, axis2 = Year, y=Total.Population))+ scale_x_discrete(limits = c("Occupation", "Gender"), expand = c(.02, .02)) +    geom_alluvium(aes(fill = Gender )) +
               geom_stratum() +
               geom_text(stat = "stratum", aes(label = after_stat(stratum)), size=2.5) +   
               theme_minimal() +
               ggtitle(" Total Male & Female workers involved in occupations") 
              
plt2
```

#conclusion:
We can notice the male-female ratio during the 5 year is same like more female connected with Customer service representative and flight attendants.

## Task 5
# Problem Statement
Stacked Bar plot to display avg wages in each PUMS.occupation identifying male and female employees using color in year 2018

```{r fig, fig.width=10 }
wj <- wages %>%
  filter(wages$Year=="2018")


wj1 <- ggplot(wj,aes(x=PUMS.Occupation,y=Average.Wage,fill=Gender)) +
  geom_bar(stat="identity")  +
   theme(axis.text.x = element_text(angle=90, hjust = 1)) + 
   labs(y="Average wage", x = "Occupation", color="Gender") +labs(title="Stacked barchart to identify avg wage of Male and Female employees in 2018 in each occupation")

wj1
```

#conclusion:
Average wages of male and female in every category of occupation is approximate same.

## Task 6
# Problem Statement:
Create a Tree Map for the Major.occupation.Group with each rectangle proportional to number of people working in detailed Occupation in 20184

```{r fig, fig.width=10}
occupation<- read.csv('occupations.csv',na.strings=(""))
head(occupation)
colnames(occupation)

library(treemapify)
count(occupation, Minor.Occupation.Group)

data <- occupation %>%
 filter(Year==2018) %>%
  group_by(Major.Occupation.Group,Detailed.Occupation) %>%
  summarise(total= Total.Population, .groups = NULL)

tree_map <- ggplot(data, aes(area=total, label=Detailed.Occupation, fill=Major.Occupation.Group, subgroup=Major.Occupation.Group)) +
           geom_treemap() +
           ggtitle("Workforce Distribution by Detailed Occupation for 2018") +
           geom_treemap_text(fontface = "italic", colour = "white", place = "centre", grow = TRUE)

tree_map
```

# Conclusion:
Most of the employees are into service occupations. Most of them happen to be waiters& waitresses , Cooks. Also, the data set might be containing only occupations from food industry. 

## Task7
# Problem Statement - Explore Plotly in R here and create any chart of your choice from any of the datasets provided in this homework. 

```{r}
hist(occupation$ID.Year)
```

```{r}
library(plotly)

df_air <- read.csv('airlines_delay.csv', na.strings = "")
req_columns <- c("carrier","year","arr_delay")

df_airline_required <- df_air[,req_columns] %>%
mutate(across(!carrier, ~replace_na(.x, 0)))

df_airline_top <- aggregate(df_airline_required$arr_delay, by=list(carrier=df_airline_required$carrier), FUN = sum) %>%
  arrange(desc(x)) 

df_airline_top_ten <- head(df_airline_top,10)

df_aggregate_year <- df_airline_required %>%
group_by(year,carrier) %>%
summarise(Total = sum(arr_delay)) 

df_airline_combined <- inner_join(df_aggregate_year, df_airline_top_ten, by = "carrier")
fig <- plot_ly(data = df_airline_combined, x = ~year, y = ~Total, color = ~carrier,mode='line')

fig
```
```{r}
# Create test data.
data <- data.frame(
  Major_occupation_group=c("Production, Transportation, & Material Moving Occupations", "Management, business, science, & arts occupations", "Service Occupations","Natural resources, construction, & maintenance occupations", "Sales & Office Occupations"),
  Number_of_People=c(105,223,130,40,148)
)
 
# Compute percentages
data$fraction = data$Number_of_People / sum(data$Number_of_People)

# Compute the cumulative percentages (top of each rectangle)
data$ymax = cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin = c(0, head(data$ymax, n=-1))
 
# Make the plot
ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Major_occupation_group)) +
     geom_rect()+coord_polar(theta="y")  + # Try to remove that to understand how the chart is built initially
     xlim(c(2, 4)) + # Try to remove that to see how to make a pie chart  
     scale_fill_manual(values = c("#B2182B","#F4A582", "#D1E5F0", "#92C5DE", "#2166AC"))
```
